R Notebook

### 2024-02-09 Scripts/authoritarian_only.Rmd
Changes to appendices and fixes in main text

In this paper, I argue that\textemdash in keeping with the idea of 'informational autocracy' (ADD CITES HERE)\textemdash an election outcome produces a signal of the relative strength of the incumbent. As Magaloni (2008) and others have noted, incumbents typically work to produce a higher margin of victory in the election, in order to demonstrate their strength and induce cooperation from other actors in society. In some theories, this only works if the votes for the incumbent are largely 'genuine' (i.e. not fabricated), or produced using the incumbent's organizational resources (i.e. vote-buying and intimidation). (ADD CITES HERE, INCLUDING TO HARVEY AND M). These two kinds of votes correspond to two of the three 'pillars of stability' identified by Gerschewski (2013) (ADD CITE): legitimation and cooptation. Fraud is considered a form of cheap talk in these models, which opposition actors can discount if they become aware of it.

However, these models overlook that engineering widespread fraud requires control over patronage resources and the criminal and judicial apparatus (ADD CITES to HARVEY). As a result, they overlook the fact that high levels of fraud are a sign of incumbent strength in both cooptation and Gerschewski's third pillar: repression. Both of these tools can be useful for defusing or dismantling an emergent protest movement. As a result, effective fraudsters should also be effective dismantlers of protest. This helps explain the broad pattern, where most fraudulent elections go unprotested, and among those that do generate protest, the vast majority fail to dislodge the incumbent. Consequently, in this view, there is little penalty for severe manipulation in terms of protest risk, and little deterrent effect.

@@

Combining data from NELDA [@hyde_2012_marinov], V-Dem [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019],... [MOVE]

@@

[REWRITE LATER] One troubling implication of these results is that ruling parties should follow an 'all-in' strategy on election fraud, a process that entails consolidating control over patronage resources and further eroding the rule of law. More fraud is not associated with longer-lasting protest waves, but a large margin of victory is associated with a lower risk of protest initiation. These results help explain both the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory. They also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation.

@@

[MOVE TO DISCUSSION AND REWRITE] Generally speaking, protest-based models predict that manipulation should decline as opposition groups become stronger relative to incumbents. This dynamic creates a sort of virtuous cycle, wherein strong opposition groups' credible threat of protest makes elections fairer, which in turn improves the strength of the opposition--leading ultimately to the ideal of self-enforcing democracy. The alternative school of thought, among other predictions, argues that opposition strength drives front-line agent behavior more than that of leaders, and causes a shift to less dispersed, less observable kinds of manipulation [@asunka2017electoral;@Harvey2019]. Election integrity can thus remain low, even in competitive contexts. Putting the protest-deterrent logic to the test, then, helps resolve significant questions in the study of election manipulation and democratization.

  Of course, electoral manipulation is not understood to be the sole cause of protest\textemdash protests occur for all sorts of reasons\textemdash but the literature has drawn a close connection between the risk of specifically electoral protest and the conduct of elections.

@@

plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c( "winner.share.upd", "v2elirreg.inv [-.05, 2]"), show.data = T)
plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()

@@

These results are consistent with different specifications presented in an online appendix. In particular, the main results are supported when data from the Mass Mobilization dataset are used in place of the ECAV data (appendix page 1), and when a measure of overall election integrity is used in place of the measure of election fraud (appendix page 15).

@@

```{r, include=FALSE}
ggplot(data=nelda2.sub, aes(x=v2elirreg.inv, y = jitter(any.ecav.protest), color = as.factor(v2x_regime))) +
  geom_point()

Predicting post-election protest counts and getting residuals:

```{r, include=FALSE}
test3 <- glm.nb(n.events.post ~  jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag +  + osorg.1lag  + natlpartyorg.lag +   inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag, data = nelda2.sub)
summary(test3)

ggplot(mapping = aes(x=predict(test3, type = "response"), y= test3$model$n.events.post)) + geom_point() + coord_cartesian(xlim=c(0, 500))

newdata <- modelr::add_predictions(nelda2.sub, test3)
newdata <- newdata %>% mutate(pred_error = n.events.post - pred)
  #because this is set up so that the predicted value is subtracted from the actual, underpredicting
  # protest is a positrive value (i.e. actual 5 - 0 predicted = 5)

test4 <- lm(pred_error ~ winner.share.upd + v2elirreg.inv + presidential, data = newdata)
summary(test4)

test5 <- lm(pred_error ~ winner.share.upd*v2elirreg.inv + presidential, data = newdata)
summary(test5)

sjPlot::plot_model(test5, type = "pred", terms = c("v2elirreg.inv", "winner.share.upd [.4, .7]"))

plot(newdata$v2elirreg.inv, newdata$pred_error)
```

Predicting the level of election manipulation, and then seeing if actual values above prediction are associated with protest

```{r, include=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(inc.confident.num = ifelse(inc.confident == "N/A", NA, ifelse(inc.confident == "yes", 1, 0)))

###For testing missingness
subsample <- nelda2.sub %>% dplyr::select(v2elirreg.inv, jucon.1lag, legcon.1lag, gdpgro.1lag,  natlpartyorg.lag,  presidential, urban.pct, rentierstate,  taxstate,  unscheduled_election, regional_success_lag, physinteg.1lag, antisys.1lag, csenv.1lag, preelection_concerns, inc.termlimit.num, liberal_no_jucon, v2elirreg.inv.lag, patronage.appts.1lag,
                                                  state.owned.1lag)
summary(subsample)
############################ Note: Drop legcon from analysis because of high missingness and high cor with lib_no_jucon

test6 <- lmer(v2elirreg.inv ~ jucon.1lag +  gdpgro.1lag +  natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + physinteg.1lag + antisys.1lag + csenv.1lag + preelection_concerns + inc.termlimit.num   + liberal_no_jucon + v2elirreg.inv.lag +
                patronage.appts.1lag +  state.owned.1lag +  v2elembaut + (1 | stateid), data = nelda2.sub) 
summary(test6)

testdata <- broom.mixed::augment(test6)
  #This adds fitted values and residuals.
  #Residuals are actual values minus predicted value, so
  #over-predicting fraud will result in a negative number (actual fraud < predicted )
  #Under-predicting fraud--what we are interested in--will result in a positive number (actual > predicted)

## Need to add rownames to nelda2.sub, so that can merge variables from nelda2.sub into testdata
nelda2.sub <- nelda2.sub %>% mutate(.rownames = rownames(nelda2.sub))
neldavars <- nelda2.sub %>% dplyr::select(.rownames, n.events.post, inc.lose, winner.share.upd, country, year, diem_explicheat_max, diem_fraud_max, country_id_year, surv_status,  inc.replaced.prot, d.events.num,
                                          legcon.1lag, osorg.1lag, altinf.1lag, libdem.1lag,  winner.share.upd, n.events.pre)
testdata <- testdata %>% left_join(neldavars, by = ".rownames")

model_protest <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd + .resid  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + .resid*winner.share.upd,
                               ziformula=~.,
                              data = testdata %>% filter(inc.lose == 0),
                                family=truncated_nbinom2)
summary(model_protest)

model_protest2 <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd + .resid  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                               ziformula=~.,
                              data = testdata %>% filter(inc.lose == 0),
                                family=truncated_nbinom2)
summary(model_protest2)

plot(testdata$winner.share.upd, testdata$.resid)

testdata <- testdata %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))
testdata <- as.data.frame(testdata)
model_protest_binary <- glm(no_protest ~ .resid*winner.share.upd, family = binomial(link = "logit"),
                             data = testdata %>% filter(inc.lose == 0))
plot_binary_margin_typ_n <- sjPlot::plot_model(model_protest_binary, type = "pred", terms = c(".resid", "winner.share.upd [.2, .8]"))
plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Excess election fraud", y = "Predicted probability", title = "Predicted probability of no protest", color = "Incumbent vote-share") + theme_bw()

model_protest2_binary <- glm(no_protest ~ .resid + winner.share.upd, family = binomial(link = "logit"),
                             data = testdata %>% filter(inc.lose == 0))

sjPlot::plot_model(model_protest2_binary, type = "pred", terms = c(".resid", "winner.share.upd [.2, .8]"), show.data = T)
sjPlot::plot_model(model_protest2_binary, type = "est", vline.color = "black") + theme_bw()

### Seeing how residuals compare to DIEM data

summary(lm(.resid ~ diem_explicheat_max + diem_fraud_max, data = testdata))
plot(testdata$diem_explicheat_max, testdata$.resid)
plot(testdata$diem_fraud_max, testdata$.resid)
  ##This shows that the resids are basically noise with regard to election observer reports, which should be good; I think this helps push back against the idea that V-Dem coders are giving higher scores to protest elections. But it would be good to push deeper and look at the correlation only among the most up-coded resids.

high_resids <- testdata %>% filter(.resid >= .222)
summary(lm(.resid ~ diem_explicheat_max + diem_fraud_max, data = high_resids))

ggplot(data = testdata, aes(x = jitter(diem_fraud_max), y = .resid)) + geom_point(aes(size = n.events.post))

plot(high_resids$diem_explicheat_max, high_resids$.resid)
plot(high_resids$diem_fraud_max, high_resids$.resid)
  #We do see more outliers here, good to think deeper--there are two major outliers that are also big protest events

### Duration

cox1_b <- coxph(Surv(d.events.num, surv_status) ~ winner.share.upd  + .resid, data = testdata %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
summary(cox1_b)

## DIEM and protest

model_protest_diem <- glmmTMB(n.events.post ~   winner.share.upd +  diem_explicheat_max ,
                               ziformula=~.,
                              data = nelda2.sub %>% filter(inc.lose == 0),
                                family=truncated_nbinom2)
summary(model_protest_diem)

```{r, include=FALSE}
nelda2.sub <- nelda2.sub %>% group_by(country) %>% mutate(elirreg_mean = mean(v2elirreg.inv, na.rm=T))
nelda2.sub <- nelda2.sub %>% mutate(elirreg_mean_diff = v2elirreg.inv - elirreg_mean)

model_protest1 <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd*elirreg_mean_diff  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                               ziformula=~.,
                              data =nelda2.sub %>% filter(inc.lose == 0),
                                family=truncated_nbinom2)
summary(model_protest1)
sjPlot::plot_model(model_protest1, type ="pred", terms = c("elirreg_mean_diff" , "winner.share.upd [.2, .8]")) + coord_cartesian(ylim = c(0, 50))

nelda2.sub <- nelda2.sub %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))
model_protest_binary <- glm(no_protest ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd*elirreg_mean_diff  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                             data =nelda2.sub %>% filter(inc.lose == 0),
                                family=binomial(link = "logit"))
summary(model_protest_binary)
sjPlot::plot_model(model_protest_binary, type="pred", terms = c("elirreg_mean_diff" , "winner.share.upd [.2, .8]"))

model_protest2 <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd + elirreg_mean_diff  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                               ziformula=~.,
                              data =nelda2.sub %>% filter(inc.lose == 0),
                                family=truncated_nbinom2)
summary(model_protest2)

Ruling party turnover

```{r, include=FALSE}
model.margin.turnover <- glm(inc.lose ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +   + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + state.owned.1lag + patronage.appts.1lag, family = binomial(link = "logit"), data = nelda2.sub)
summary(model.margin.turnover)

m.inclose <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag*patronage.appts.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub %>% filter(inc.lose == 1)) 
summary(m.inclose)
sjPlot::plot_model(m.inclose, type = "int")

m.incwin <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag*patronage.appts.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub %>% filter(inc.lose == 0)) 
summary(m.incwin)
sjPlot::plot_model(m.incwin, type = "int")

### Z score

z <- (-1.647988 - (-1.144192)) / sqrt((0.691932)^2 + (0.329801)^2) #B1 - B2 / sqrt((seB1)^2 + (seB2)^2)

>>>>

In this paper, I argue \textemdash in keeping with a growing literature on the informational aspects of elections in authoritarian regimes [@guriev2020;@smyth2020elections;@Geddes1989;@Miller2015a]\textemdash that an election outcome produces a signal of the relative strength of the incumbent. As Magaloni [-@magaloni2006] nd others have noted, incumbents typically work to produce a higher margin of victory in the election, in order to demonstrate their strength and induce cooperation from other actors in society. In some theories, this only works if the votes for the incumbent are largely 'genuine' (i.e. not fabricated), or produced using the incumbent's organizational resources (i.e. vote-buying and intimidation) [@Harvey2020;@magaloni2006]. These two kinds of votes correspond to two of the three 'pillars of stability' identified by Gerschewski [-@Gerschewski2013]: legitimation and cooptation. Fraud is considered a form of cheap talk in these models, which opposition actors can discount if they become aware of it.

However, these models overlook that engineering widespread fraud requires control over patronage resources and the criminal and judicial apparatus [@Harvey2019;@Harvey2022]. As a result, they overlook the fact that high levels of fraud are a sign of incumbent strength in both cooptation and Gerschewski's third pillar: repression. Both of these tools can be useful for defusing or dismantling an emergent protest movement. As a result, effective fraudsters should also be effective dismantlers of protest. This helps explain the broad pattern, where most fraudulent elections go unprotested, and among those that do generate protest, the vast majority fail to dislodge the incumbent. Consequently, in this view, there is little penalty for severe manipulation in terms of protest risk, and little deterrent effect.

@@

@@

One troubling implication of these results is that\textemdash if protest is indeed costly to incumbents\textemdash ruling parties should follow an 'all-in' strategy on election fraud, a process that entails consolidating control over patronage resources and further eroding the rule of law. More fraud is not associated with longer-lasting protest waves, but a large margin of victory is associated with a lower risk of protest initiation. These results help explain several empirical patterns at once: the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory in conjunction with high levels of fraud. At a theoretical elvel, they also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation. Finally, they suggest a pessimistic outlook for democratization by election in authoritarian regimes.

@@

@@

plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c( "winner.share.upd", "v2elirreg.inv [-.05, 2]"), show.data = T)
plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()

@@

These results are consistent with different specifications presented in an online appendix. In particular, the main results are supported when data from the Mass Mobilization dataset are used in place of the ECAV data (appendix page 1), and when a measure of overall election integrity is used in place of the measure of election fraud (appendix page 15). [CHECK PAGES]

[MOVE TO DISCUSSION AND REWRITE] Generally speaking, protest-based models predict that manipulation should decline as opposition groups become stronger relative to incumbents. This dynamic creates a sort of virtuous cycle, wherein strong opposition groups' credible threat of protest makes elections fairer, which in turn improves the strength of the opposition--leading ultimately to the ideal of self-enforcing democracy. The alternative school of thought, among other predictions, argues that opposition strength drives front-line agent behavior more than that of leaders, and causes a shift to less dispersed, less observable kinds of manipulation [@asunka2017electoral;@Harvey2019]. Election integrity can thus remain low, even in competitive contexts.

@@
